---
title: "UK - EDA"
author: "Author-Name"
date: "5/21/2021"
output: html_document
---


```{r message=FALSE}
mypack <- function(package){
  new.package <- package[!(package %in% installed.packages()[, "Package"])]
  if (length(new.package)) 
    install.packages(new.package, dependencies = T)
  sapply(package, require, character.only = T)
}
packages <- c("dplyr", "caret", "tidyr", "tibble", "forcats", "knitr", "corrplot",
              "esquisse", "fastDummies", "plotly", "lubridate", "timeDate",
              "gridExtra", "visdat", "IRdisplay", "highcharter")
mypack(packages)

```

### Load cleaned file
```{r, message=FALSE}

data_accidents <- read.csv(file.choose(), header=T)
data_accidents <- data_accidents %>% rename(`Severity_accident`=`Accident_Severity`,
                                            `Rural_Urban` = `Urban_or_Rural_Area`)
data_accidents <- as_tibble(data_accidents)
accidents_numerical_df <- dummy_cols(data_accidents,
                                     select_columns = c('Region', 'Rural_Urban', 'X1st_Road_Class', 'Road_Type',
                                                        'Road_Surface_Conditions', 'Weather', 'High_Wind', 'Lights',
                                                        'Junction_Detail', 'Junction_Location', 'X1st_Point_of_Impact',
                                                        'Driver_Journey_Purpose', 'Propulsion_Code', 'Vehicle_Make',
                                                        'Vehicle_Category', 'Vehicle_Manoeuvre'),
                                     remove_most_frequent_dummy=TRUE)

columns_to_drop <- c('Accident_Index', 'Latitude', 'Longitude', 'Datetime', 'Region', 'Rural_Urban',
                     'X1st_Road_Class', 'Road_Type', 'Road_Surface_Conditions', 'Weather', 'High_Wind',
                     'Lights', 'Junction_Detail', 'Junction_Location', 'X1st_Point_of_Impact',
                     'Driver_Journey_Purpose', 'Propulsion_Code', 'Vehicle_Make', 'Vehicle_Category',
                     'Vehicle_Manoeuvre')

accidents_numerical_df <- accidents_numerical_df[ , !(names(accidents_numerical_df) %in% columns_to_drop)]
accidents_numerical_df_normalized <- predict(preProcess(accidents_numerical_df, method=c("center", "scale")), accidents_numerical_df)
accidents_numerical_df_normalized <- dummy_cols(accidents_numerical_df_normalized,
                                                select_columns = c('Severity_accident'),
                                                remove_most_frequent_dummy=TRUE)

```


### Data Sample
```{r echo=FALSE}
columns_to_drop <- c('Severity_accident')
names(accidents_numerical_df_normalized)[names(accidents_numerical_df_normalized) == 'Severity_accident_Fatal_Serious'] <- 'Fatal_or_Serious_Accident'
accidents_numerical_df_normalized <- accidents_numerical_df_normalized[ , !(names(accidents_numerical_df_normalized) %in% columns_to_drop)]
accidents_numerical_df <- dummy_cols(accidents_numerical_df,
                                     select_columns = c('Fatal_or_Serious_Accident'),
                                     remove_most_frequent_dummy=TRUE)

str(accidents_numerical_df)
```



```{r echo=FALSE}
columns_to_drop <- c('Severity_accident')
names(accidents_numerical_df)[names(accidents_numerical_df) == 'Severity_accident_Fatal_Serious'] <- 'Fatal_or_Serious_Accident'
accidents_numerical_df <- accidents_numerical_df[ , !(names(accidents_numerical_df) %in% columns_to_drop)]
rm(columns_to_drop)
```


### Correlation plots with the corrplot package
```{r echo=FALSE}
cols_to_filter <- c('Driver_IMD_Decile', 'Speed_limit', 'Year', 'Season', 'Month_of_Year',
                    'Day_of_Month', 'Day_of_Week', 'Hour_of_Day', 'Number_of_Vehicles',
                    'Age_of_Driver', 'Age_of_Vehicle', 'Engine_CC', 'Fatal_or_Serious_Accident')
corrplot(cor(accidents_numerical_df_normalized[, cols_to_filter]), type="upper", method="circle", tl.col="black", tl.cex=0.6, tl.offset=0.5)



#3


```



```{r echo=FALSE}
cols_to_filter <- c('Rural_Urban_Rural', 'X1st_Road_Class_B', 'X1st_Road_Class_Unclassified',
                    'X1st_Road_Class_C', 'X1st_Road_Class_Motorway', 
                    'X1st_Road_Class_A(M)', 'Road_Type_Roundabout', 'Road_Type_Dual carriageway',
                    'Road_Type_One way street', 'Road_Type_Slip road', 
                    'Road_Surface_Conditions_Wet or damp', 'Road_Surface_Conditions_Frost or ice',
                    'Road_Surface_Conditions_Snow', 
                    'Road_Surface_Conditions_Flood over 3cm. deep', 'High_Wind_Yes', 'Propulsion_Code_Heavy oil',
                    'Fatal_or_Serious_Accident')
corrplot(cor(accidents_numerical_df_normalized[, cols_to_filter]), type="upper", method="circle", tl.col="black",
         tl.cex=0.6, tl.offset=0.5)
```



```{r echo=FALSE}
cols_to_filter <- c('Junction_Detail_Crossroads', 'Junction_Detail_T or staggered junction',
                    'Junction_Detail_More than 4 arms (not roundabout)',
                    'Junction_Detail_Roundabout', 'Junction_Detail_Private drive or entrance', 'Junction_Detail_Slip road',
                    'Junction_Detail_Other junction', 'Junction_Location_Mid Junction - on roundabout or on main road',
                    'Junction_Location_Approaching junction or waiting/parked at junction approach',
                    'Junction_Location_Cleared junction or waiting/parked at junction exit',
                    'Junction_Location_Leaving main road',
                    'Junction_Location_Entering main road', 'Junction_Location_Leaving roundabout',
                    'Junction_Location_Entering from slip road',
                    'Junction_Location_Entering roundabout', 'X1st_Point_of_Impact_Offside',
                    'X1st_Point_of_Impact_Nearside',
                    'X1st_Point_of_Impact_Back', 'X1st_Point_of_Impact_Did not impact', 'Fatal_or_Serious_Accident')
corrplot(cor(accidents_numerical_df_normalized[, cols_to_filter]), type="upper", method="circle", tl.col="black", tl.cex=0.6, tl.offset=0.5)
```
### Histogram 
#### Histogram Severity
```{r echo=FALSE}

p1 <- plot_ly(alpha = 0.6) %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Slight", ]$Season, histnorm="probability", name="Slight") %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Fatal_Serious", ]$Season, histnorm="probability", name="Fatal-Serious") %>%
  layout(barmode="overlay", title="Season comparison", xaxis=list(title="Season"))
print(p1)

```

#### Histogram age vs severity
```{r echo=FALSE}

p2 <- plot_ly(alpha = 0.6) %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Slight", ]$Age_of_Driver, histnorm="probability", name="Slight") %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Fatal_Serious", ]$Age_of_Driver, histnorm="probability", name="Fatal-Serious") %>%
  layout(barmode="overlay", title="Age of Driver comparison", xaxis=list(title="Age of Driver"))
print(p2)
```

#### Histogram weather vs severity
```{r echo=FALSE}
p4 <- plot_ly(alpha = 0.6) %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Slight", ]$Weather, histnorm="probability", name="Slight") %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Fatal_Serious", ]$Weather, histnorm="probability", name="Fatal-Serious") %>%
  layout(barmode="overlay", title="Weather comparison", xaxis=list(title="Weather"))
print(p4)

```

#### Histogram region vs severity
```{r echo=FALSE}
p5 <- plot_ly(alpha = 0.6) %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Slight", ]$Region, histnorm="probability", name="Slight") %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Fatal_Serious", ]$Region, histnorm="probability", name="Fatal-Serious") %>%
  layout(barmode="overlay", title="Region comparison", xaxis=list(title="Region"))
print(p5)

```

#### Histogram ruralurban vs severity
```{r echo=FALSE}

p6 <- plot_ly(alpha = 0.6) %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Slight", ]$Rural_Urban, histnorm="probability", name="Slight") %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Fatal_Serious", ]$Rural_Urban, histnorm="probability", name="Fatal-Serious") %>%
  layout(barmode="overlay", title="Urban or Rural Area comparison", xaxis=list(title="Urban or Rural Area"))
print(p6)

```

#### Speedlimit vs severity
```{r echo=FALSE}

p7 <- plot_ly(alpha = 0.6) %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Slight", ]$Speed_limit, histnorm="probability", name="Slight") %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Fatal_Serious", ]$Speed_limit, histnorm="probability", name="Fatal-Serious") %>%
  layout(barmode="overlay", title="Speed Limit comparison", xaxis=list(title="Speed Limit"))
print(p7)
```

#### Road type
```{r echo=FALSE}

p8 <- plot_ly(alpha = 0.6) %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Slight", ]$Road_Type, histnorm="probability", name="Slight") %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Fatal_Serious", ]$Road_Type, histnorm="probability", name="Fatal-Serious") %>%
  layout(barmode="overlay", title="Road Type comparison", xaxis=list(title="Road Type"))
print(p8)



```

#### Road surface
```{r echo=FALSE}
p9 <- plot_ly(alpha = 0.6) %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Slight", ]$Road_Surface_Conditions, histnorm="probability", name="Slight") %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Fatal_Serious", ]$Road_Surface_Conditions, histnorm="probability", name="Fatal-Serious") %>%
  layout(barmode="overlay", title="Road Surface Conditions comparison", xaxis=list(title="Road Surface Condition"))
print(p9)

```

#### Lights comparison
```{r echo=FALSE}
p10 <- plot_ly(alpha = 0.6) %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Slight", ]$Lights, histnorm="probability", name="Slight") %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Fatal_Serious", ]$Lights, histnorm="probability", name="Fatal-Serious") %>%
  layout(barmode="overlay", title="Lights comparison", xaxis=list(title="Light"))
print(p10)


```

#### Junction comparison
```{r echo=FALSE}

p11 <- plot_ly(alpha = 0.6) %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Slight", ]$Junction_Detail, histnorm="probability", name="Slight") %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Fatal_Serious", ]$Junction_Detail, histnorm="probability", name="Fatal-Serious") %>%
  layout(barmode="overlay", title="Junction comparison", xaxis=list(title="Junction"))
print(p11)


```

#### 1st point Impact
```{r echo=FALSE}
p12 <- plot_ly(alpha = 0.6) %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Slight", ]$X1st_Point_of_Impact, histnorm="probability", name="Slight") %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Fatal_Serious", ]$X1st_Point_of_Impact, histnorm="probability", name="Fatal-Serious") %>%
  layout(barmode="overlay", title="1st Point of Impact comparison", xaxis=list(title="1st Point of Impact"))
print(p12)


```

#### Vehicle Make
```{r echo=FALSE}

p13 <- plot_ly(alpha = 0.6) %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Slight", ]$Vehicle_Make, histnorm="probability", name="Slight") %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Fatal_Serious", ]$Vehicle_Make, histnorm="probability", name="Fatal-Serious") %>%
  layout(barmode="overlay", title="Vehicle Make comparison", xaxis=list(title="Vehicle Make"))
print(p13)


```

#### Vehicle category
```{r echo=FALSE}

p14 <- plot_ly(alpha = 0.6) %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Slight", ]$Vehicle_Category, histnorm="probability", name="Slight") %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Fatal_Serious", ]$Vehicle_Category, histnorm="probability", name="Fatal-Serious") %>%
  layout(barmode="overlay", title="Vehicle Category comparison", xaxis=list(title="Vehicle Category"))
print(p14)

```

#### Hours day vs severity
```{r echo=FALSE}
p16 <- plot_ly(alpha = 0.6) %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Slight", ]$Hour_of_Day, histnorm="probability", name="Slight", nbinsx=100) %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Fatal_Serious", ]$Hour_of_Day, histnorm="probability", name="Fatal-Serious", nbinsx=100) %>%
  layout(barmode="overlay", title="Hour of Day comparison", xaxis=list(title="Hour of Day"))
print(p16)

```

#### Days vs severity
```{r echo=FALSE}
p17 <- plot_ly(alpha = 0.6) %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Slight", ]$Day_of_Week, histnorm="probability", name="Slight") %>%
  add_histogram(x=~data_accidents[data_accidents$Severity_accident == "Fatal_Serious", ]$Day_of_Week, histnorm="probability", name="Fatal-Serious") %>%
  layout(barmode="overlay", title="Day of Week comparison", xaxis=list(title="Day of Week"))
print(p17)
```


### Heatmap hour day week
#### All Accidents
```{r echo=FALSE}


p1 <- select(data_accidents, one_of(c("Accident_Index", "Datetime", "Severity_accident"))) %>%
  mutate(dhour = format(strptime(Datetime,"%Y-%m-%d %H:%M:%S"),'%H'),
         wday = wday(Datetime, label=TRUE)) %>%
  mutate(wday = fct_relevel(wday, c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))) %>%
  group_by(dhour, wday) %>%
  summarise(accidents_count = n_distinct(Accident_Index))  %>%
  ggplot(aes(dhour, wday, fill=accidents_count)) +
  geom_tile() +
  coord_fixed() +
  labs(x="Hour of Day", y="Day of Week") +
  scale_fill_distiller(palette="Spectral", name="Accidents Count") +
  ggtitle("All Accidents") +
  theme(plot.margin=grid::unit(c(2,2,2,2), "mm"))

# Fatal-Serious Accidents
p2 <- select(data_accidents, one_of(c("Accident_Index", "Datetime", "Severity_accident"))) %>%
  filter(Severity_accident=="Fatal_Serious") %>%
  mutate(dhour = format(strptime(Datetime,"%Y-%m-%d %H:%M:%S"),'%H'),
         wday = wday(Datetime, label=TRUE)) %>%
  mutate(wday = fct_relevel(wday, c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))) %>%
  group_by(dhour, wday) %>%
  summarise(accidents_count = n_distinct(Accident_Index))  %>%
  ggplot(aes(dhour, wday, fill=accidents_count)) +
  geom_tile() +
  coord_fixed() +
  labs(x="Hour of Day", y="Day of Week") +
  scale_fill_distiller(palette="Spectral", name="Accidents Count") +
  ggtitle("Fatal-Serious Accidents") +
  theme(plot.margin=grid::unit(c(2,2,2,2), "mm"))

# Slight Accidents
p3 <- select(data_accidents, one_of(c("Accident_Index", "Datetime", "Severity_accident"))) %>%
  filter(Severity_accident=="Slight") %>%
  mutate(dhour = format(strptime(Datetime,"%Y-%m-%d %H:%M:%S"),'%H'),
         wday = wday(Datetime, label=TRUE)) %>%
  mutate(wday = fct_relevel(wday, c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))) %>%
  group_by(dhour, wday) %>%
  summarise(accidents_count = n_distinct(Accident_Index))  %>%
  ggplot(aes(dhour, wday, fill=accidents_count)) +
  geom_tile() +
  coord_fixed() +
  labs(x="Hour of Day", y="Day of Week") +
  scale_fill_distiller(palette="Spectral", name="Accidents Count") +
  ggtitle("Slight Accidents") +
  theme(plot.margin=grid::unit(c(2,2,2,2), "mm"))

grid.arrange(p1, p2, p3, ncol=1)

```
